\begin{thebibliography}{200}

% A stick-slip motivacioja
\vspace*{-0.3cm}
\bibitem{Spider1} F.\ Vollrath \'es D.\ Porter,\ Soft Matter\ {\bf2} (2006) 377.

\vspace*{-0.3cm}
\bibitem{Spider2} T.\ A.\ Blackledge, A.\ P.\ Summers \'es C.\ Y.\ Hayashi,\ Zoology {\bf 108}, (2005) 41-46.

\vspace*{-0.3cm}
\bibitem{Spider3} S. Frische, A. B. Maunsbach \'es F. Vollrath,\ journal of Microscopy {\bf 189}, (1998) 64-70.

\vspace*{-0.3cm}
\bibitem{Spider4} F. K. Ko, S. Kawabata, M. Inoue, M. Niwa, S. Fossey \'es J. W. Song, Material Research Society, Fall Meeting (2001).

\vspace*{-0.3cm}
\bibitem{Spider5} B. Madsen, z. Z. Shao \'es F. Vollrath, International Journal of Biological Macromolecules {\bf 24}, (1999) 301-306.

\vspace*{-0.3cm}
\bibitem{Violin1} J.\ Ringlein,\ Science Teacher\ {\bf72}, (2005) 24-32.

\vspace*{-0.3cm}
\bibitem{Violin2} Forr\'as: {\it www.phys.unsw.edu.an/jw/Bows.html}.

\vspace*{-0.3cm}
\bibitem{Trib1} M.\ Muraki,\ E.\ Kinbaru \'es T.\ Konishi,\ Tribology International\ {\bf36}, (2003) 739-744.

\vspace*{-0.3cm}
\bibitem{Trib2} W.\ S.\ Owen \'es E.\ A.\ Croft,\  IEEE/ASME Transactions on Mechatronics\ {\bf8}, (2003) 362-371.

\vspace*{-0.3cm}
\bibitem{Comp1} R.\ Joseph, M.\ T.\ Martyn \'es P.\ D.\ Coates, Bulletin of Material Science\ {\bf 29}, {\it Springer India}, (2006) 85-89.

\vspace*{-0.3cm}
\bibitem{Comp2} J.\ Karger-Kocsis, Journal of Macromechanical Science, Part:B,\ {\bf 40} (2001) 343-353.

\vspace*{-0.3cm}
\bibitem{SS1} R.\ C.\ Hidalgo, F.\ Kun, K.\ Kov\'acs \'es I.\ Pagonabarraga, Physical Review E {\bf80}, (2009) 051108.

\vspace*{-0.3cm}
\bibitem{SS2} F.\ Raischel, F.\ Kun \'es H.\ J.\ Herrmann, Physical Review E {\bf64}, (2001) 066122.

\vspace*{-0.3cm}
\bibitem{DIF1} J.\ P.\ Sethna, K.\ A.\ Dahmen \'es C.\ R.\ Meyers, Nature {\bf410}, (2001) 2422.

\vspace*{-0.3cm}
\bibitem{DIF2} F-J.\ Perez-Reche, L.\ Truskinovsky \'es G.\ Zanotto, Advances in Physics {\bf101}, (2008) 230601.

\vspace*{-0.3cm}
\bibitem{DIF3} J.\ P.\ Sethna, K.\ A.\ Dahmen, S.\ Kartha, B.\ W.\ Roberts \'es J.\ L.\ Shoren, Physical Review Letters {\bf70}, (1993) 3347.

\vspace*{-0.3cm}
\bibitem{DIF4} J.\ V.\ Andersen, D.\ Sornette \'es K.\ Leung, Physical Review Letters {\bf78}, (1997) 2140.


%FBM

\vspace*{-0.3cm}
\bibitem{FBM_Pei} F.\ T.\ Peires, Journal of Textil Institute\ {\bf 17}, (1926) T355.

\vspace*{-0.3cm}
\bibitem{FBM_Dani} H.\ E.\ Daniels, Proceedings of Royal Society, London\ A {\bf 183}, (1945) 192.

\vspace*{-0.3cm}
\bibitem{FBM_Cole} B.\ D.\ Coleman, Journal of Applied Physics {\bf 27}, (1956) 862.

\vspace*{-0.3cm}
\bibitem{FBM_std1} S.\ L.\ Phoenix \'es B.\ I.\ J., Statistical Strength Theory for Fibrious Composite Materials, a {\it Comprehensive Composite Materials} k\"{o}tetben, A.\ Kelley \'es C.\ Zweben szerkeszt\'es\'eben {\bf 1}, 1.19 Fejezet, Pergamon-Elsevier Science, New York, 2000.

\vspace*{-0.3cm}
\bibitem{FBM_std2} B.\ K.\ Chakrabarti \'es L. G. Benguigui, {\it Statistical Physics of Fracture and Disordered Systems}, Oxford University Press, 1997.

\vspace*{-0.3cm}
\bibitem{FBM_std3} H.\ J.\ Herrmann \'es S.\ Roux, editors, {\it Statistical models for the fracture of disordered media}, Random materials and Processes, Elsevier-Amsterdam, 1990.

\vspace*{-0.3cm}
\bibitem{FBM_std4} J.\ V.\ Andersen, D.\ Sornette \'es K.\ Leung, Physical Review Letters\ {\bf 78}, (1997) 2140.

\vspace*{-0.3cm}
\bibitem{FBM_std5} M.\ Kloster, A.\ Hansen \'es P.\ C.\ Hemmer, Physical Review E {\bf 56}, (1997) 2615.

\vspace*{-0.3cm}
\bibitem{FBM_std6} S.\ Pradhan \'es B.\ K.\ Chakrabarti, International Journal of modern Physics B {\bf 17}, (2003) 5565.

\vspace*{-0.3cm}
\bibitem{FBM_LLS1} J.\ B.\ G\'omez, D. I\~{n}iguez \'es A.\ F.\ Pacheco, Physical Review Letters {\bf 71}, (1993) 380.

\vspace*{-0.3cm}
\bibitem{FBM_LLS2} W.\ A.\ Curtin, Physical Review Letters {\bf 80}, (1998) 1445.

\vspace*{-0.3cm}
\bibitem{FBM_LLS3} A.\ Hansen \'es P.\ C. Hemmer, Physical Letters A {\bf 184}, (1994) 394.

\vspace*{-0.3cm}
\bibitem{FBM_LLS4} D.\ G.\ Harlow, Proceedings of Royal Society, London A {\bf 397}, (1985) 211.

\vspace*{-0.3cm}
\bibitem{FBM_LLS5} R.\ C.\ Hidalgo, Y.\ Moreno, F.\ Kun \'es H.\ J.\ Herrmann Physical Review E {\bf 65}, (2002) 046148.

\vspace*{-0.3cm}
\bibitem{FBM_LLS6} G.\ Dill-Langer \'es m\'asok, Physica A {\bf 235}, (2003) 547.

\vspace*{-0.3cm}
\bibitem{FBM_LLS7} F.\ Kun, Y.\ Moreno, R.\ C.\ Hidalgo \'es H.\ J.\ Herrmann, Europhysics Letters {\bf 63}, (2003) 347.

\vspace*{-0.3cm}
\bibitem{FBM_Ava1} A.\ Hansen and S.\ Roux, Statistical toolbox for damage and fracture, a {\it Damage and Fracture of Disordered Materials} kötetben, D. Krajnoviocic \'es J.\ van Mier szerkeszt\'es\'eben, CISM Courses and Lectures No. 410, 2. Fejezet, 17-101, Springer Verlag, 2000.

\vspace*{-0.3cm}
\bibitem{FBM_Ava2} S.\ Pradhan , A.\ Hansen \'es P.\ C.\ Hemmer, Physical Review Letters {\bf 95}, (2005) 125501.

% Bevezet\'es
\vspace*{-0.3cm}
\bibitem{Leo} L.\ da\ Vinci, {\it I libri di Meccanica}, Milano, 1940.

\vspace*{-0.3cm} 
\bibitem{Intro1} M.\ Janssen, J.\ Zuidema \'es R. Wanhill, {\it Fracture Mechanics}, SPON Publishing, 2004.

\vspace*{-0.3cm} 
\bibitem{Intro2} T.\ L.\ Anderson, {\it Fracture Mechanics: Fundamentals and Applications}, Taylor \& Francis Group, 2005.

\vspace*{-0.3cm} 
\bibitem{Intro3} M.\ J.\ Alava, P.\ K.\ N.\ N.\ Nukala \'es S.\ Zapperi, Advances in Physics {\bf 55}, (2006) 349-476.

\vspace*{-0.3cm}
\bibitem{Intro4} V.\ L.\ Emery \'es M.\ Morgenstein, Journal of Archaeological Science {\bf 34}, (2007).

\vspace*{-0.3cm}
\bibitem{Intro5} A.\ Lucas \'es J.\ R.\ Harris: Ancient Egyptian Materials and Industries 1962, reprinted by Histories and Mysteries of Man LTD. London, 1989.

\vspace*{-0.3cm}
\bibitem{Intro6} L.\ A.\ Sawyer \'es W.\ H.\ Mitchell Liberty Ships: The History of the Emergency Type Cargo Ships Constructed in the United States During the Second World War, Lloyd's of London Press Ltd, 1985.

\vspace*{-0.3cm}
\bibitem{Intro7} X.\ Wu, X-Y.\ Liu, N.\ Du, G.\ Xu \'es B-W.\ Li, Molecular spring: from spider silk to silkworm silk, (2009).

%stat fiz

\vspace*{-0.3cm}
\bibitem{Stat1} W.\ Weibull, Proceedings of Royal Swedish Institute Of Engineering Research {\bf 151}, (1939).

\vspace*{-0.3cm}
\bibitem{Stat2} R.\ Furth, Nature {\bf 145}, (1940), 7413.

\vspace*{-0.3cm}
\bibitem{Stat3} Z.\ Jaeger \'es R.\ Englman, Thermodynamical Theory for Fracture in Heterogeneous Solids a {\it Damage Mechanics in Engineering Materials} k\"{o}tetben, New York 1989.

\vspace*{-0.3cm}
\bibitem{Stat4} B.\ T.\ Brady, A Thermodynamic Basis for Static and Dynamic Scaling Laws in the Design of Structures in Rock, a {\it Rock Mechanics}, k\"{o}tetben by P.\ Nelson \'es S.\ Laubach szerkeszt\'es\'eben, Balkema, Rotterdam, 1994.

\vspace*{-0.3cm}
\bibitem{Stat5} M. Ostoja-Starzewski, Damage in Random Microstructure: Size Effects, Fractals and Entropy Maximization, in {\it Mechanics Pan-America}, edited by C.\ R.\ Steele, A.\ W.\ Leissa \'es M.\ R.\ M.\ Crespo da Silva, ASME Press, New York, 1989.

\vspace*{-0.3cm}
\bibitem{Stat6} S.\ N.\ Zhurkov, International Journal of Fracture Mechanics {\bf 1}, 19653.2.

\vspace*{-0.3cm}
\bibitem{Stat7} T.\ L.\ Chelidze, Pure and Applied Geophysics {\bf 124}, (1986) 731-748.

\vspace*{-0.3cm}
\bibitem{Stat8} A.\ Delaplace , G.\ Pijaudier-Cabot \'es S.\ Roux, International Journal of Mechanical Physics of Solids, {\bf 44}, (1999) 99-106.

\vspace*{-0.3cm}
\bibitem{Stat9} C.\ W.\ Nan, Progress in Materials Science {\bf 37}, (1993) 1163.

% timedep
\vspace*{-0.3cm}
\bibitem{TDI1} W.\ A.\ J.\ Albert, Archive f\"{u}r Mineralogie Geognosie Bergbau und H\H{u}ttenkunde {\bf 10}, (1838) 215-234.

\vspace*{-0.3cm}
\bibitem{TDI2} W.\ J.\ M.\ Rankine, Institution of Civil Engineers, Minutes of Proceedings, (1842) 105-108.

\vspace*{-0.3cm}
\bibitem{TDI3} J.\ Glynn, Rudimentary treatise on the construction of cranes and machinery, High Holborn, 1954.

\vspace*{-0.3cm}
\bibitem{TDI3_1} A.\ Wöhler, Zeitschrift f\"{u}r Bauwesen {\bf 5} (1855) 121-166.

\vspace*{-0.3cm}
\bibitem{TDI3_2} O.\ H.\ Basquin, Proceedings of American Society of Testing Materials ASTEA {\bf 10}, (1910) 625.

\vspace*{-0.3cm}
\bibitem{TDI3_3} A.\ G.\ Palmgren, Zeitschrift des Vereines Deutscher Ingenieure (VDI Zeitschrift) {\bf 68}, (1924) 339-341.

\vspace*{-0.3cm}
\bibitem{book_pap} K. Niskanen, I. Kajanto \'es P. Pekarinen, Paper structure a {\it Paper Physics (Book 16) from Papermaking Science and Technology} k\"{o}tetben, J. Gullischen \'es H. Paulapuro szerkeszt\'es\'eben, Finnish Paper Engineer's Association and Technical Association of the Pulp and Paper Industry, 1998.

\vspace*{-0.3cm}
\bibitem{TDI4} P.\ C.\ Kersavage, Wood Fiber Science. {\bf 5(2)}, 1(973) 105-117.

\vspace*{-0.3cm}
\bibitem{TDI5} R.\ Wath\`{e}n, Studies on fiber strength and its effect on paper properties, PhD disszert\'aci\'o, Helsinki University of Technology, Espoo, 2006.

\vspace*{-0.3cm}
\bibitem{TDI6} D.\ H.\ Page \'es F.\ El-Hosseiny, Pulp and Paper in Canada {\bf 84(9)}, (1983) TR99.

\vspace*{-0.3cm}
\bibitem{TDI7} K.\ W.\ Hardacker, Technical Association of the Pulp and Paper Industry {\bf 8}, (1970) 201-216.

\vspace*{-0.3cm}
\bibitem{TDI8} B.\ A.\ Jayne, Technical Association of the Pulp and Paper Industry {\bf 42(6)}, (1957) 461.

\vspace*{-0.3cm}
\bibitem{TDI9} C.\ A.\ Jentzen, Technical Association of the Pulp and Paper Industry {\bf 47(7)}, (1964) 112.

\vspace*{-0.3cm}
\bibitem{TDI10}  M.\ T.\ Kortschot, Fracture of paper a {\it Handbook of Physical Testing of Paper} {\bf 1}. kötete 2. kiad\'as\'aban, R.\ E.\ Mark, C.\ C.\ Habeger Jr., J.\  Borch, M.\ B.\ Lyne szerkeszt\'es\'eben, Marcel Decker, New York, 2002.

\vspace*{-0.3cm}
\bibitem{TDI11} T. Yamauchi, S. Okumura \'es N. Noguchi, Journal of Pulp and Paper Science {\bf 16}, (1990).

\vspace*{-0.3cm}
\bibitem{PP1} Global Forest, Paper \& Packaging Industry Survey 2008 Edition, PricewaterhouseCoopers, 2008. 

\vspace*{-0.3cm}
\bibitem{PP2} Forr\'as: {\it www.unesco.org}.

\vspace*{-0.3cm}
\bibitem{CRE1} F.\ Kun, Z.\ Hal\'asz, J.\ S.\ Andrade \'es H.\ J.\ Herrmann, Journal of Statistical Mechanics: Theory and Experiment, (2008).

\vspace*{-0.3cm}
\bibitem{CRE2} F.\ Kun, H.\ A.\ Carmona, J.\ S.\ Andrade \'es H.\ J.\ Herrmann, Physical Review Letters {\bf 100}, (2008) 094301.

\vspace*{-0.3cm}
\bibitem{CRE3} F.\ Kun, R.\ Cruz Hidalgo, H.\ J.\ Herrmann \'es K.\ F.\ Pal, Physical Review E {\bf 67}, (2003), 061802.

\vspace*{-0.3cm}
\bibitem{CRE4} I.\ G.\ Main, Geophysical Journal International {\bf 142}, (2000) 151.

\vspace*{-0.3cm}
\bibitem{CRE5} T.\ Baxevanis \'es T. Katsaounis, Europhysics Letters {\bf 81}, (2008) 24001.

\vspace*{-0.3cm}
\bibitem{CRE6} T.\ Baxevanis \'es T. Katsaounis, Eoropean Physical Journal B {\bf 61}, (2008) 153.

\end{thebibliography}
